On Pure Exploration, Thresholding Bandits, and Kullback-Leibler Divergence

### 2017-09-18 Thesis/Draft.tex
Printed version

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={On Pure Exploration, Thresholding Bandits, and Kullback-Leibler Divergence},
            colorlinks=true,
            linkcolor=Maroon,
            citecolor=Blue,
            urlcolor=blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setlength{\droptitle}{-2em}
  \title{On Pure Exploration, Thresholding Bandits, and Kullback-Leibler
Divergence}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

@@

\begin{document}
\maketitle

{
\hypersetup{linkcolor=black}
\setcounter{tocdepth}{2}
\tableofcontents
}
\newpage

@@

``In general, we look for a new law by the following process: First we
guess it. Then we -- now don't laugh, that's really true. Then we
compute the consequences of the guess to see what, if this is right, if
this law that we guessed is right, to see what it would imply. And then
we compare the computation results to nature, or we say compare to
experiment or experience, compare it directly with observations to see
if it works. If it disagrees with experiment, it's wrong. In that simple
statement is the key to science. It doesn't make any difference how
beautiful your guess is, it doesn't make any difference how smart you
are, who made the guess, or what his name is. If it disagrees with
experiment, it's wrong. That's all there is to it.''

\begin{itemize}
\tightlist
\item
  Richard Feynman in \emph{Richard Feynman Messenger Lectures at Cornell
  University: The Character of Physical Law. Chapter 7, Seeking New
  Laws}.
\end{itemize}

@@

@@

This makes sequential sampling especially relevant in situations where
there is an inherent cost to obtaining samples, or where only a limited
amount of samples can be collected. The classical example is that of
clinical trials in which a researcher tests two drugs on patients to
observe which one performs better. Not only is it difficult to gather
enough patients, but one would also like to adapt the sampling such that
as soon as possible more patients receive the better drug and are cured.
As noted in Cappé et al. (2013), this example dates back to Thompson
(1933, 1935). Such problems are known as multi-armed bandit problems in
which the sampling is adapted so as to minimize the cumulative regret:
In the case of the drug tests, one would like to sample the two drugs in
a way to find the best drug as soon as possible and then to test this
drug more and more often and more often than the worse drug. This would
lead to cumulatively more healthy patients than testing both drugs
uniformly; it would minimize the regret from treating patients with an
ineffective drug.

Instead of drugs, one can consider two or more unknown distributions
(which we call arms in the multi-armed bandit problem in analogy with
the one-armed bandit slot machine). These distributions have different
means, and one can draw samples sequentially from the distributions in
order to estimate their means. To minimize the cumulative regret
measure, one would like to play the distribution with the highest mean
as much as possible (\emph{Exploitation}). At the same time, however,
one has to sample all arms at least a little, so as to initially find
the best arm (\emph{Exploration}). Any algorithm thus has to handle a
trade-off between exploration and exploitation to minimize the regret.
How can we organize the sampling in order to achieve this task?

While Gittins (1979) presents an Bayesian-optimal solution to the
cumulative regret problem, the Upper-Confidence Bound (UCB) algorithms
pioneered by Agrawal (1995) and Auer, Cesa-Bianchi and Fischer (2002)
have lead to a blooming literature on multi-armed bandit problems that
has been picked up to solve modern problems in website and advertising
optimization, where advertisers try to allocate their advertising
optimally so as to reduce regret from showing poorly converting
advertisements. Several algorithms like the KL-UCB algorithm based on
Kullback-Leibler divergence upper confidence intervals (Cappé et al.,
2013) or the Bayes-UCB algorithm (Kaufmann et al., 2012) have followed
and been shown to be optimal as they reach the lower bound introduced in
Lai and Robbins (1985) and Burnetas and Katehakis (1996).

@@

Similar algorithms based on indexes can be used on related problems.
Instead of minimizing cumulative regret, it can be of interest to find
the best arm, that is, the distribution with the largest mean -- with as
few samples as possible and with confidence as large as possible. This
is particularly useful when there is not a cost from testing arms with
lower mean, but when drawing any sample is costly. Thus one needs to
explore the distributions as efficiently as possible. Exploitation is
not necessary, and the setting is thus referred to as \emph{Pure
Exploration}. The need for strategies other than uniform sampling
arises: One does not want to continue to sample distributions that one
is confident in. This setting is related to the problem faced in
Bayesian Optimization: Training large neural networks is costly in terms
of time and computation, and so the parameter space needs to be explored
with as few samples as possible. (Zhang et al., 2017; Brochu et al.,
2010). The problem of best arm identification is also very related to
statistical hypothesis testing as noted by Chernoff (1958) since one has
to make a decision about how the means of distributions relate to each
other. The main difference being that the bandit setting includes the
testing procedure in the sampling process.

Given the focus on exploration, the pure exploration problem is usually
described with a loss function different from the one in the cumulative
regret settings. Audibert et al. (2010) introduce a soft measure called
\emph{simple regret} based on which they describe UCB policies. Other
authors investigate binary measures that are equal to 1 when the best
arm has not been identified (Russo, 2016; Garivier and Kaufmann, 2016).
These binary measures correspond to the probability of the identified
arm being the best, and thus inherently related to hypothesis testing.
The problem of best arm identification can readily be extended to the so
called Top-\(m\) problem of identifying the best \(m\) of \(K\) arms
(Kaufmann and Kalyanakrishnan, 2013; Kaufmann et al. 2016).

In this thesis, we consider the thresholding bandit problem. While the
name is reminiscient of the cumulative regret multi-armed bandit, the
problem is more similar to the top-\(m\) problem. But instead of
identifying the best \(m\) out of \(K\) arms, one aims to find all
distributions out of \(K\) that have a mean at least as large as a fixed
threshold. Given that an implicit threshold arises in the top-\(m\)
problem between the \(m\)-th best and the \((m+1)\)-th best arm
(Kaufmann et al., 2016), the two problems share many characteristics. We
discuss this in more detail in
\autoref{chap:KullbackLeiblerPureExploration}. In the form described in
Locatelli et al. (2016), algorithms in the thresholding bandit problem
are evaluated against a binary loss function penalizing whenever an arm
has not been classified correctly as being above or below the threshold,
and thus focus on exploration similar to algorithms employed for best
arm identification. Indeed, Locatelli et al. (2016) show how their
thresholding algorithm can be used for best arm identification when the
mean of the best arm is known upfront.

@@

Consider the problem the latter companies face. Whenever a user reports
that she visited a restaurant, the app would like to ask questions about
her experience. To give a simple example, assume users are not willing
to answer more than one question per restaurant visit, and the app would
like to find out two things: Does the restaurant have good service? Is
the restaurant trendy?

@@

The app will then display a ``Good service'' or ``This place is trendy''
description for the restaurant if \(\mu_{\text{service}}\) and
\(\mu_{\text{trendy}}\) are estimated to be above a certain threshold,
for example \(\tau = 0.5\). That is, more than half of the people would
describe the place as having good service, or as trendy.

To make the app as useful as possible to other users, this information
should be shown as quickly as possible. But even though the app has
thousands of users, not that many people visit a specific restaurant,
and even less people report this back to the app. Thus the company only
has a limited budget of \(T\) potential users from which to collect
samples \(X_{\text{service}}\) and \(X_{\text{trendy}}\) from. How can
the company ask these \(T\) questions to be as confident as possible in
its classification, that is, in the information it displays?

A straightforward way of sampling users would be to ask the two
questions in a round-robin fashion: Ask for ``Service'', then for
``Trendy'', then for ``Service'', \ldots{} Given 100 users, 50 users
give feedback on the service level, 50 users report on the trendiness
level. This however is unlikely to be optimal with respect to the
confidence after 100 samples. Given that the company asks the users
sequentially, there is the option to adapt the sampling strategy over
time: If the first ten users report
\(\sum_{s=1}^{10}X_{\text{service},s}=10\), while
\(\sum_{s=1}^{10}X_{\text{trendy},s}=5\), then the company might already
be quite confident that \(\mu_{\text{service}} \geq \tau = 0.5\), while
this is less clear for \(\mu_{\text{trendy}}\). Instead of continuously
sampling both questions, it might be more useful to collect more
feedback on the trendiness level so as to figure out its position
relative to the threshold. After all, the company would like to use the
100 samples to maximize its confidence in both displayed tips.

Finding the optimal sequential sampling strategy to classify
distributions against a threshold with as much confidence as possible
after a limited budget of samples is the problem described above, and
the problem described by the thresholding bandit problem which we
discuss for the remainder of this thesis. In the following section, we
define the general setting of the thresholding bandit problem.

@@

Consider a setting with a set of \(\mathbb{A}=\{1, ..., K\}\) unknown
probability distributions \(\nu_i, i \in \mathbb{A}\). To stick with the
idea of multi-armed bandits, we call the distributions arms, and
observations from the arms rewards (Bubeck and Cesa-Bianchi, 2012). The
learner is interested in learning about the structure of the arms; in
particular, she would like to estimate the means \(\mu_1, ..., \mu_K\)
by sequentially drawing observations \(X_{i,1}, X_{i,2}, ...\) from the
different arms, where \(X_{i} \sim \nu_i\). However, at time \(t\) the
learner cannot observe a reward from every distribution but must select
a single arm \(I_t\) from which she would like to observe the reward
\(X_{I_t,t}\). Often, adaptive sampling strategies makes the decision of
which arm to pull by computing an index \(B_i(t)\) for each arm. If
\(T_i(t)\) is the number of samples observed until time \(t\) for arm
\(i\), then is \(B_i(t)\) a function of the past rewards
\(X_{i,1}, ..., X_{i,T_i(t)}\) observed for arm \(i\), and w.l.o.g.
\(I_t = \arg \min_{i = 1,...K} B_i(t)\). The set of indexes can be
called strategy (Kaufmann et al., 2012) which makes it clear that the
strategy is adapted as more observations are collected. Strategies of
this kind are also called index policies.

@@

The thresholding problem for the learner is now to use a budget of
overall \(T\) samples (or rounds) to explore the different arms. After
this exploration phase, the learner returns a set of arms that she
believes to be over the threshold. When evaluating the classification of
the learner, arms with mean inside of the epsilon interval
\(\tau \pm \epsilon\) will not be considered. The learner only has to
correctly distinguish the means from the threshold up to the precision
given by \(\epsilon\). Thus, the algorithm is not evaluated against
\(\mathcal{S}_{\tau}\) and \(\mathcal{S}^C_{\tau}\), but needs to
distinguish \(\mathcal{S}_{\tau+\epsilon}\) from
\(\mathcal{S}^C_{\tau-\epsilon}\).

The learner's classification is returned as the set
\(\hat{\mathcal{S}}_{\tau} := \hat{\mathcal{S}}_{\tau}(T) \subset \mathbb{A}\)
after \(T\) rounds; that is, the set of arms estimated to be over the
threshold. The classification is evaluated against the loss given as
(Locatelli et al., 2016):

@@

Algorithms for this thresholding problem can be compared based on their
expected loss. Let \(\mathbb{E}\) be the expectation with regard to the
samples \(X_{i,1}, ..., X_{i,T}\) collected by the algorithm. The
expected loss is then given as

@@

which corresponds to the probability of making at least one wrong
classification. Both the probability of not identifying an arm that is
actually over the threshold, as well as the probability of accepting an
arm as over the threshold when it is not, are evaluated. The oracle
strategy would evaluate all arms correctly with an error probability of
\(0\) (Locatelli et al., 2016).

At this point, we have introduced the general thresholding bandit
problem setting which we will use throughout the thesis. The setting can
be extended by different assumptions to make certain algorithms or
bounds feasible. For that reason, Locatelli et al. (2016) assume that
the distributions \(\nu_i, i \in \mathbb{A}\) are \(R\)-sub-Gaussian.
Under this assumption, the authors are able to prove a lower bound on
the largest probability of error of the best algorithm in the setting we
have just introduced, as well as an upper bound on the performance of
their APT algorithm which we will introduce in the following chapter. As
in their paper, we define \(R\)-sub-Gaussian and assume it in the
description of the their algorithm and bounds:

@@

As noted by Locatelli et al. (2016), this ensures that the mean
\(\mu_i\) of the distributions \(\nu_i\) considered in our problem are
finite. \(R\)-sub-Gaussian distributions offer concentration
inequalities for the concentration of the empirical mean around the true
mean, which makes it possible to deduce the bounds presented by the
authors and repeated below. The family of \(R\)-sub-Gaussian
distributions encompasses Gaussian distributions of variance \(R^2\) for
\(R \in \mathbb{R}\), as well as all bounded distributions, with the
important case of Bernoulli distributions.

Having described the thresholding bandit setting as well as the
assumption of \(R\)-sub-Gaussianity, we can now proceed to describe the
\emph{Anytime Parameter-free Thresholding} (APT) algorithm proposed in
Locatelli et al. (2016) for the thresholding bandit problem.

@@

\begin{algorithm}
\SetKwData{Left}{left}\SetKwData{This}{this}\SetKwData{Up}{up}
\SetKwFunction{Union}{Union}\SetKwFunction{FindCompress}{FindCompress}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{$\tau$, $\epsilon$}
\Output{$\hat{S}_{\tau} = \{i \in \{1,...,K\}, \hat{\mu}_i(T) \geq \tau\}$}
\BlankLine
\emph{Pull each arm once}\;
\For{$t = K+1$ \KwTo $T$}{
  \emph{Pull arm $I_t = \arg \min_{i \in A} B_i^{APT}(t) = \arg \min_{i \in A} \sqrt{T_i(t)} (|\hat{\mu}_i(t) - \tau|+\epsilon)$}\;
  \emph{Observe reward $X_t \sim \nu_{I_t}$}\;
}
\caption{Anytime Parameter-free Thresholding Algorithm (Locatelli et al., 2016)}\label{alg:APT}
\end{algorithm}

@@

As noted in Zhong et al. (2017), a simple requirement to correctly
classify arms at the final round is the following. For every arm
\(i \in \{1, ..., K\}\) after the final round \(T\), require \[
\hat{\mu}_{i}(T) \geq \mu_i - \frac{\mu_i - \tau + \epsilon}{2} \quad \text{if} \quad \mu_i \geq \tau
\] or \[
\hat{\mu}_{i}(T) \leq \mu_i + \frac{\tau - \mu_i + \epsilon}{2} \quad \text{if} \quad \mu_i < \tau
\] Or simply in terms of the gap between the arm and the threshold as
used in the index by the APT algorithm: \[
| \hat{\mu}_{i}(T) - \mu_i | \leq \frac{|\mu_i - \tau| + \epsilon}{2} = \frac{\Delta_i}{2}.
\]

>>>>

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            colorlinks=true,
            linkcolor=black,
            citecolor=Blue,
            urlcolor=black,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

@@

\begin{document}

\pagenumbering{gobble}

\begin{centering}

\vspace{5cm}

    {\Large Humboldt-Universit\"at zu Berlin \\
    School of Business and Economics \\
    Institute for Statistics and Econometrics\\}

    \vspace{4cm}

{\huge{\bf On Pure Exploration, Thresholding Bandits, \\
and Kullback-Leibler Divergence\\}}

\vspace{2cm}

{\large Master's Thesis submitted\\\vspace{0.5cm}
    to}
    \\\vspace{0.5cm}
    {\large{\bf Prof. Dr. Marius Kloft \\}}
    {\large{\bf Prof. Alexandra Carpentier, PhD \\}}
    \vspace{0.5cm}

    {\large by \\\vspace{0.5cm}
    {\bf Tim Radtke\\}
    (572481)\\
    born on 11.01.1992 in Viersen \\\vspace{1cm}}

    {\large in partial fulfillment of the requirements \\
    for the degree of \\
    {\bf Master of Science (M. Sc.)\\} \vspace{3cm}
    Berlin, September 18, 2017\\}

\end{centering}

@@

``In general, we look for a new law by the following process: First we
guess it. Then we -- now don't laugh, that's really true. Then we
compute the consequences of the guess to see what, if this is right, if
this law that we guessed is right, to see what it would imply. And then
we compare the computation results to nature, or we say compare to
experiment or experience, compare it directly with observations to see
if it works. If it disagrees with experiment, it's wrong. In that simple
statement is the key to science. It doesn't make any difference how
beautiful your guess is, it doesn't make any difference how smart you
are, who made the guess, or what his name is. If it disagrees with
experiment, it's wrong. That's all there is to it.''
\vspace*{1.5\baselineskip}

Richard Feynman in \emph{Richard Feynman Messenger Lectures at Cornell
University: The Character of Physical Law. Chapter 7, Seeking New Laws}.

@@

\Large{\bf Abstract\\}

\vspace*{1\baselineskip}

\normalsize
In this thesis we present a new sequential sampling strategy for the
fixed-budget thresholding bandit problem described in Locatelli et al.
(2016). We draw from the literature on pure exploration bandit problems
to show the similarity of the thresholding bandit to (sequential)
hypothesis testing. In particular, we derive a lower bound on how fast
the error probability in a one-armed thresholding bandit problem can
decrease at best, and show that this corresponds to the Chernoff-Stein
Lemma. This motivates the use of the Kullback-Leibler divergence in the
index of our new algorithm. In simulations, our strategy is shown to
outperform existing strategies. Using website analytics data from an
online shop, we use time series cross-validation to create data sets for
offline testing of adaptive sampling strategies; time dependency in the
data is thus preserved. Our experiments show that the adaptive sampling
strategies are disrupted by the dependency structure in the samples.

\setcounter{tocdepth}{2} \tableofcontents

\pagenumbering{arabic}

@@

This makes sequential sampling especially relevant in situations where
there is an inherent cost to obtaining samples, or where only a limited
amount of samples can be collected. The classical example is that of
clinical trials in which a researcher tests two drugs on patients to
observe which one performs better. Not only is it difficult to gather
enough patients, but one would also like to adapt the sampling such that
more patients receive the better drug as soon as possible, and are
cured. As noted in Cappé et al. (2013), this example dates back to
Thompson (1933, 1935). Such problems are known as multi-armed bandit
problems in which the sampling is adapted so as to minimize the
cumulative regret: In the case of the drug tests, one would like to
sample the two drugs in a way that quickly finds the best drug and then
tests this drug more and more often, and more often than the worse drug.
This would lead to cumulatively more healthy patients than testing both
drugs uniformly; it would minimize the regret from treating patients
with an ineffective drug.

Instead of drugs, one can consider two or more unknown distributions
(which we call arms in the multi-armed bandit problem in analogy with
the one-armed bandit slot machine). These distributions have different
means, and one can draw samples sequentially from the distributions in
order to estimate their means. To minimize the cumulative regret
measure, one would like to sample from the distribution with the highest
mean as often as possible (\emph{Exploitation}). At the same time,
however, one has to sample all arms at least a little, so as to
initially find the best arm (\emph{Exploration}). Any algorithm thus has
to handle a trade-off between exploration and exploitation to minimize
the regret. How can we organize the sampling in order to achieve this
task?

While Gittins (1979) presents an Bayesian-optimal solution to the
cumulative regret problem, the Upper-Confidence Bound (UCB) algorithms
pioneered by Agrawal (1995) and Auer, Cesa-Bianchi and Fischer (2002)
have lead to a blooming literature on multi-armed bandit problems that
has been picked up to solve modern problems in website and advertising
optimization, where advertisers try to allocate their advertising
optimally so as to reduce regret from showing poorly converting
advertisements. Several algorithms like the KL-UCB algorithm based on
Kullback-Leibler divergence upper confidence intervals (Cappé et al.,
2013) or the Bayes-UCB algorithm (Kaufmann et al., 2012a) have followed
and been shown to be optimal as they reach the lower bound introduced in
Lai and Robbins (1985) and Burnetas and Katehakis (1996).

@@

Similar algorithms based on indexes can be used on related problems.
Instead of minimizing cumulative regret, it can be of interest to find
the best arm, that is, the distribution with the largest mean -- with as
few samples as possible and with confidence as large as possible. This
is particularly useful when there is not a cost from testing arms with
lower mean, but when drawing any sample is costly. Thus one needs to
explore the distributions as efficiently as possible. Given that the
loss function only evaluates the classifiation once at the end of the
sampling process, exploitation is \emph{not} necessary, and the setting
is thus referred to as \emph{Pure Exploration}. The need for strategies
different from uniform sampling arises still: One does not want to
continue to sample distributions that one is already confident in. This
setting is related to the problem faced in Bayesian Optimization:
Training of, for example, deep neural network architectures is costly in
terms of time and computation, and so the parameter space needs to be
explored with as few samples as possible. (Zhang et al., 2017; Brochu et
al., 2010). The problem of best arm identification is also very related
to sequential statistical hypothesis testing as noted by Chernoff
(1958), and picked up by Garivier and Kaufmann (2016) and Russo (2016),
since the experimenter has to make a decision about how the means of
distributions relate to each other.

Given the focus on exploration, the pure exploration problem is usually
described with a loss function different from the one in the cumulative
regret settings. Audibert et al. (2010) introduce a soft measure called
\emph{simple regret} based on which they describe UCB policies. Other
authors investigate binary measures that are equal to 1 when the best
arm has not been identified (Russo, 2016; Garivier and Kaufmann, 2016).
These binary measures correspond to the probability of the identified
arm being the best, and thus are inherently related to hypothesis
testing. The problem of best arm identification can readily be extended
to the so called Top-\(m\) problem of identifying the best \(m\) of
\(K\) arms (Kaufmann and Kalyanakrishnan, 2013; Kaufmann et al. 2016).

In this thesis, we consider the thresholding bandit problem. While the
name is reminiscient of the cumulative regret multi-armed bandit, the
problem is more similar to the top-\(m\) problem. But instead of
identifying the best \(m\) out of \(K\) arms, one aims to find all
distributions out of \(K\) that have a mean at least as large as a fixed
threshold (it is not known upfront how many distributions fulfill this
condition). Given that an implicit threshold arises in the top-\(m\)
problem between the \(m\)-th best and the \((m+1)\)-th best arm
(Kaufmann et al., 2016), the two problems share many characteristics. We
discuss this in more detail in
\autoref{chap:KullbackLeiblerPureExploration}. In the setting described
in Locatelli et al. (2016), algorithms in the thresholding bandit
problem are evaluated against a binary loss function penalizing once
after the sampling process if an arm has not been classified correctly
as being above or below the threshold. This setting thus focuses on
exploration similar to algorithms employed for best arm identification.
Indeed, Locatelli et al. (2016) show how their thresholding algorithm
can be used for best arm identification when the mean of the best arm is
known upfront.

@@

Consider the problem the latter companies face. Whenever a user reports
that she visited a restaurant, the app would like to ask questions about
her experience. To give a simple example, assume users are not willing
to answer more than one question per restaurant visit, and the app would
like to find out two things: Does the restaurant have good service? Is
the restaurant trendy? See \autoref{app:Foursquare} for how this
information is then displayed in the Foursquare app.

@@

The app will display a ``Good service'' or ``This place is trendy''
description for the restaurant if \(\mu_{\text{service}}\) and
\(\mu_{\text{trendy}}\) are estimated to be above a certain threshold,
for example \(\tau = 0.5\). That is, more than half of the user base
would describe the place as having good service, or as trendy.

To make the app as useful as possible to other users, this information
should be shown as quickly as possible. But even though the app has
thousands of users, not that many people visit a specific restaurant,
and even less people report this back to the app. Thus the company has
only a limited budget of \(T\) potential users from which to collect
samples \(X_{\text{service}}\) and \(X_{\text{trendy}}\). How can the
company ask these \(T\) questions to be as confident as possible in its
classification of the restaurant's features, that is, in the information
it displays?

A straightforward way of sampling users would be to ask the two
questions in a round-robin fashion: Ask for ``Service'', then for
``Trendy'', then for ``Service'', \ldots{} Given 100 users, 50 users
give feedback on the service level, 50 users report on the trendiness
level. This, however, is unlikely to be optimal with respect to the
confidence after 100 samples. Given that the company asks the users
sequentially, there is the option to adapt the sampling strategy over
time: If the first ten users report
\(\sum_{s=1}^{5}X_{\text{service},s}=5\), while
\(\sum_{s=1}^{5}X_{\text{trendy},s}=3\), then the company might already
be quite confident that \(\mu_{\text{service}} \geq \tau = 0.5\), while
this is less clear for \(\mu_{\text{trendy}}\). Instead of continuously
sampling both questions, it might be more useful to collect more
feedback on the trendiness level so as to ensure its position relative
to the threshold. After all, the company would like to use the 100
samples to maximize its confidence in \emph{both} displayed tips.

Finding the optimal sequential sampling strategy to classify
distributions against a threshold with as much confidence as possible
after a limited budget of samples is the problem described above. It is
a setting described by the thresholding bandit problem, which we discuss
for the remainder of this thesis. In the following section, we define
setting of the thresholding bandit problem more generally.

@@

Consider a setting with unknown probability distributions \(\nu_i\)
indexed by \(i \in \mathbb{A}=\{1, ..., K\}\). To stick with the idea of
multi-armed bandits, we call the distributions arms, and observations
from the arms rewards (Bubeck and Cesa-Bianchi, 2012). The learner is
interested in learning about the structure of the arms; in particular,
it would like to estimate the means \(\mu_1, ..., \mu_K\) of each
distribution \(\nu_i\) by sequentially drawing observations
\(X_{i,1}, X_{i,2}, ...\) from the different arms, where
\(X_{i} \sim \nu_i\). However, at time \(t\) the learner cannot observe
a reward from every distribution but must select a single arm \(I_t\)
from which it would like to observe the reward \(X_{I_t,t}\). The
decision of which arm to pull next is often made by computing an index
\(B_i(t)\) for each arm. If \(T_i(t)\) is the number of samples observed
until time \(t\) for arm \(i\), then \(B_i(t)\) is a function of the
past rewards \(X_{i,1}, ..., X_{i,T_i(t)}\) observed for arm \(i\), and
w.l.o.g. \(I_t = \arg \min_{i = 1,...K} B_i(t)\). The set of indexes can
be called strategy (Kaufmann et al., 2012a) which makes it clear that
the strategy is adapted as more observations are collected. Strategies
of this kind are also called index policies.

@@

The thresholding problem for the learner is now to use a budget of
overall \(T\) samples (or rounds) to explore the different arms. After
this exploration phase, the learner returns a set of arms that it
believes to be over the threshold. When evaluating the classification of
the learner, arms with mean inside of the epsilon interval
\(\tau \pm \epsilon\) are not considered. The learner only has to
correctly distinguish the means from the threshold up to the precision
given by \(\epsilon\). Thus, the algorithm is not evaluated against
\(\mathcal{S}_{\tau}\) and \(\mathcal{S}^C_{\tau}\), but needs to
distinguish \(\mathcal{S}_{\tau+\epsilon}\) from
\(\mathcal{S}^C_{\tau-\epsilon}\).

The learner's classification is returned as the set
\(\hat{\mathcal{S}}_{\tau} := \hat{\mathcal{S}}_{\tau}(T) \subseteq \mathbb{A}\)
after \(T\) rounds; that is, the set of arms estimated to be over the
threshold. The classification is evaluated against the binary loss given
by (Locatelli et al., 2016):

@@

Algorithms for this thresholding problem can be compared based on their
expected loss. Let \(\mathbb{E}\) be the expectation with regard to the
samples \(X_{i,1}, ..., X_{i,T}\) collected by the algorithm. Then

@@

is the expected loss, or the probability of making at least one wrong
classification. Both the probability of not identifying an arm that is
actually over the threshold, as well as the probability of accepting an
arm as over the threshold when it is not, are evaluated. The oracle
strategy would evaluate all arms correctly with an error probability of
\(0\) (Locatelli et al., 2016).

At this point, we have introduced the general thresholding bandit
problem setting which we will use throughout the thesis. The setting can
be extended by different assumptions to make certain algorithms or
bounds feasible. For that reason, Locatelli et al. (2016) assume that
the distributions \(\nu_i, i \in \mathbb{A}\) are \(R\)-sub-Gaussian.
Under this assumption, the authors are able to prove a lower bound on
the largest probability of error of the best algorithm in the setting
just introduced, as well as an upper bound on the performance of their
APT algorithm which we will show in the following chapter. As in their
paper, we define \(R\)-sub-Gaussian as follows:

@@

As noted by Locatelli et al. (2016), this ensures that the mean
\(\mu_i\) of the distributions \(\nu_i\) considered in our problem are
finite. \(R\)-sub-Gaussian distributions offer concentration
inequalities for the concentration of the empirical mean around the true
mean, which makes it possible to deduce the bounds presented by the
authors and repeated below. Among other distributions, the family of
\(R\)-sub-Gaussian distributions encompasses Gaussian distributions of
variance \(R^2\) for \(R \in \mathbb{R}\), as well as all bounded
distributions, with the important case of Bernoulli distributions.

Having described the thresholding bandit setting as well as the
assumption of \(R\)-sub-Gaussianity, we now proceed to describe the
\emph{Anytime Parameter-free Thresholding} (APT) algorithm proposed in
Locatelli et al. (2016) for the thresholding bandit problem.

@@

\begin{algorithm}
\SetKwData{Left}{left}\SetKwData{This}{this}\SetKwData{Up}{up}
\SetKwFunction{Union}{Union}\SetKwFunction{FindCompress}{FindCompress}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{$\tau$, $\epsilon$}
\Output{$\hat{S}_{\tau} = \{i \in \{1,...,K\}, \hat{\mu}_i(T) \geq \tau\}$}
\BlankLine
\emph{Pull each arm once}\;
\For{$t = K+1$ \KwTo $T$}{
  \emph{Pull arm $I_t = \arg \min_{i \in A} B_i^{APT}(t) = \arg \min_{i \in A} \sqrt{T_i(t-1)} \cdot(|\hat{\mu}_i(t-1) - \tau|+\epsilon)$}\;
  \emph{Observe reward $X_t \sim \nu_{I_t}$}\;
}
\caption{Anytime Parameter-free Thresholding Algorithm (Locatelli et al., 2016)}\label{alg:APT}
\end{algorithm}

@@

As noted in Zhong et al. (2017), a simple requirement to correctly
classify arms at the final round is the following. For every arm
\(i \in \{1, ..., K\}\) after the final round \(T\), require: \[
| \hat{\mu}_{i}(T) - \mu_i | \leq \frac{|\mu_i - \tau| + \epsilon}{2} = \frac{\Delta_i}{2}.
\]
